KidsAIR: Main analysis initial steps

### 2020-10-12 lrti_tracking.Rmd
analysis updates

lrti_no <- lrti_notes_decisions %>% 
  mutate(child_id_num = as.character(child_id_num),
         home_winter_id = as.character(home_winter_id),
         winter_id = as.character(winter_id)) %>% 
  # join lrti data with full dataset
  full_join(health_exposure_data_sampling_day, 
            by = c("area", "home", "child_id_num", "home_winter_id", "winter_id")) %>% 
  filter(is.na(lrti_final)) %>% 
  # select vars for calculating person-time
  select(area, home, child_id_num, home_winter_id, winter_id, lrti_final, 
         lrti_start_date_final, lrti_end_date_final, notes_final, total_cases_this_winter,
         csq_date) %>% 
  left_join(intervention_dates, by = c("area", "home_winter_id")) %>% 
  group_by(area, child_id_num, winter_id) %>% 
  arrange(area, child_id_num, winter_id, csq_date) %>% 
  # var for person-time start date for each winter/child combo
  # this date is intervention date + 14 days
  mutate(person_time_start = intervention_date + 14) %>% 
  filter(!is.na(csq_date)) %>% 
  # var for person-time end date for each winter/child combo
  # this date is the earliest of the last csq or april 30 of the study year
  mutate(person_time_end = last(csq_date),
         latest_date = "04-30") %>% 
  separate(person_time_end, into = c("pte_year", "pte_mo", "pte_day"), sep = "-", remove = FALSE) %>% 
  unite(latest_date, c("pte_year", "latest_date"), sep = "-") %>% 
  select(-pte_mo, -pte_day) %>% 
  mutate(latest_date = ymd(latest_date),
         person_time_end = if_else(person_time_end > latest_date,
                                   latest_date, person_time_end),
         person_time_end = if_else(person_time_end < person_time_start, last(csq_date), person_time_end),
         person_time_possible = as.numeric(person_time_end - person_time_start),
         person_time_possible = if_else(person_time_possible < 0, 0,
                                        person_time_possible)) %>% 
  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
  arrange(area, child_id_num, winter_id)

>>>>

lrti_no <- lrti_notes_decisions %>% 
  mutate(child_id_num = as.character(child_id_num),
         home_winter_id = as.character(home_winter_id),
         winter_id = as.character(winter_id)) %>% 
  # join lrti data with full dataset
  full_join(health_exposure_data_sampling_day, 
            by = c("area", "home", "child_id_num", "home_winter_id", "winter_id")) %>% 
  filter(is.na(lrti_final)) %>% 
  # select vars for calculating person-time
  select(area, home, child_id_num, home_winter_id, winter_id, lrti_final, 
         lrti_start_date_final, lrti_end_date_final, notes_final, total_cases_this_winter,
         csq_date) %>% 
  left_join(intervention_dates, by = c("area", "home_winter_id")) %>% 
  group_by(area, child_id_num, winter_id) %>% 
  arrange(area, child_id_num, winter_id, csq_date) %>% 
  # var for person-time start date for each winter/child combo
  # this date is intervention date + 14 days
  mutate(person_time_start = intervention_date + 14) %>% 
  filter(!is.na(csq_date)) %>% 
  mutate(person_time_start = if_else(is.na(person_time_start), first(csq_date) + 14, person_time_start)) %>% 
  # var for person-time end date for each winter/child combo
  # this date is the earliest of the last csq or april 30 of the study year
  mutate(person_time_end = last(csq_date),
         latest_date = "04-30") %>% 
  separate(person_time_end, into = c("pte_year", "pte_mo", "pte_day"), sep = "-", remove = FALSE) %>% 
  unite(latest_date, c("pte_year", "latest_date"), sep = "-") %>% 
  select(-pte_mo, -pte_day) %>% 
  mutate(latest_date = ymd(latest_date),
         person_time_end = if_else(person_time_end > latest_date,
                                   latest_date, person_time_end),
         person_time_end = if_else(person_time_end < person_time_start, last(csq_date), person_time_end),
         person_time_possible = as.numeric(person_time_end - person_time_start),
         person_time_possible = if_else(person_time_possible < 0, 0,
                                        person_time_possible)) %>% 
  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
  arrange(area, child_id_num, winter_id)

### 2020-10-12 main_analysis_initial.Rmd
analysis updates

lrti_final_data <- read_rds(paste0(file_path, "Output/lrti_final_data.rds")) %>% 
  arrange(area, home, child_id_num, winter_id) %>% 
  mutate(cohort = round_date(csq_date, unit = "year")) %>% 
  separate(cohort, c("cohort", "trash"), sep = "-") %>% 
  group_by(area, child_id_num, home_winter_id) %>% 
  mutate(lrti_events = n()) %>% 
  group_by(area, child_id_num) %>% 
  distinct(area, child_id_num, home_winter_id, .keep_all = TRUE) %>% 
  mutate(lrti_events = as.numeric(lrti_events),
         lrti_events = if_else(is.na(person_time_no_lrti_total), 0, lrti_events),
         lrti_events_total = sum(lrti_events, na.rm = TRUE),
         person_time_no_lrti_total = as.numeric(person_time_no_lrti_total),
         time_at_risk_total = sum(person_time_no_lrti_total, na.rm = TRUE),
         lrti_incidence = lrti_events_total/sum(time_at_risk_total)*7,
         lrti_incidence = if_else(is.na(lrti_incidence), 0, lrti_incidence)) %>% 
  distinct(area, child_id_num, .keep_all = TRUE) %>% 
  select(area, home, cohort, child_id_num, lrti_events_total, time_at_risk_total, lrti_incidence)
```

```{r}
# Merge dataset with blinded treatments and final LRTI data, select analysis variables
analysis_data <- health_exposure_data_sampling_day %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  left_join(treatments_masked, by = "home") %>% 
  right_join(lrti_final_data, by = c("area", "home", "child_id_num")) %>% 
  distinct(area, child_id_num, .keep_all = TRUE) %>% 
  select(area, home, cohort, child_id_num, lrti_events_total, time_at_risk_total, lrti_incidence, newtx) %>% 
  rename(treatment = newtx)
```

@@

```{r}
# Initial summary data for LRTI
incidence_summary <- analysis_data %>% 
  group_by(treatment) %>% 
  summarize(mean_inc = mean(lrti_incidence, na.rm = TRUE),
            sd_inc = sd(lrti_incidence, na.rm = TRUE),
            n_inc = n(),
            min_inc = min(lrti_incidence, na.rm = TRUE),
            median_inc = median(lrti_incidence, na.rm = TRUE),
            max_inc = max(lrti_incidence, na.rm = TRUE))
incidence_summary

# Histogram for distribution of data
incidence_plot <- analysis_data %>% 
  ggplot() +
    geom_histogram(aes(lrti_incidence)) +
    theme_light() +
    facet_wrap(~treatment)
incidence_plot

@@

nb_results <- glm.nb(lrti_incidence ~ treatment, data = analysis_data)
odTest(nb_results)

@@

# Link for analysis ideas and notes:
# https://stats.idre.ucla.edu/r/dae/negative-binomial-regression/ 
# https://stats.idre.ucla.edu/r/dae/poisson-regression/ 
# https://stats.idre.ucla.edu/r/faq/random-coefficient-poisson-models/ 
```

```{r}
# Mixed model
model_results <- glmer.nb(lrti_incidence ~ treatment + (1 | home:cohort:area), data = analysis_data)

summary(model_results)
```

>>>>

lrti_final_data <- read_rds(paste0(file_path, "Output/lrti_final_data.rds")) %>% 
  arrange(area, home, child_id_num, winter_id) %>% 
  mutate(cohort = round_date(csq_date, unit = "year")) %>% 
  separate(cohort, c("cohort", "trash"), sep = "-") %>% 
  group_by(area, child_id_num, winter_id) %>% 
  mutate(lrti_events = n()) %>% 
  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
  group_by(area, child_id_num) %>% 
  mutate(lrti_events = as.numeric(lrti_events),
         lrti_events = if_else(is.na(person_time_no_lrti_total), 0, lrti_events),
         lrti_events_overall = sum(lrti_events, na.rm = TRUE),
         person_time_no_lrti_total = as.numeric(person_time_no_lrti_total),
         person_time_no_lrti_total = if_else(is.na(person_time_no_lrti_total), person_time_possible,
                                             person_time_no_lrti_total),
         person_time_at_risk_total = sum(person_time_no_lrti_total, na.rm = TRUE),
         person_time_at_risk_total = as.numeric(person_time_at_risk_total)/7,
         lrti_incidence = lrti_events_overall/person_time_at_risk_total,
         lrti_incidence = if_else(is.na(lrti_incidence), 0, lrti_incidence)) %>% 
  arrange(area, home, child_id_num, winter_id) %>% 
  distinct(area, child_id_num, .keep_all = TRUE) %>% 
  ungroup() %>% 
  filter(person_time_at_risk_total > 0) %>% 
  select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence)
```

```{r}
# Merge dataset with blinded treatments and final LRTI data, select analysis variables
analysis_data <- health_exposure_data_sampling_day %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  mutate(child_age = fct_collapse(age_child_nov,
                                  "< 1" = "Less than 1",
                                  "1+" = c(1, 2, 3, 4))) %>% 
  left_join(treatments_masked, by = "home") %>% 
  arrange(area, home, child_id_num, winter_id, age_child_nov) %>% 
  right_join(lrti_final_data, by = c("area", "home", "child_id_num")) %>% 
  distinct(area, child_id_num, .keep_all = TRUE) %>% 
  select(area, home, cohort, child_id_num, child_age, lrti_events_overall, 
         person_time_at_risk_total, lrti_incidence, newtx) %>% 
  rename(treatment = newtx,
         lrti_events = lrti_events_overall,
         person_time_at_risk = person_time_at_risk_total) %>% 
  mutate(log_person_time_at_risk = log(person_time_at_risk)) %>% 
  mutate(lrti_events_di = if_else(lrti_events == 0, 0, 1)) %>% 
  filter(lrti_incidence < 0.7) %>% 
  mutate(treatment = factor(treatment,
                            levels = c("Dearborn", "Smith", "Blackfoot")))
```

@@

```{r}
# Initial summary data for LRTI
incidence_summary <- analysis_data %>% 
  group_by(treatment) %>% 
  #filter(lrti_incidence < 0.7) %>% 
  summarize(mean_inc = mean(lrti_incidence, na.rm = TRUE),
            sd_inc = sd(lrti_incidence, na.rm = TRUE),
            n_inc = n(),
            min_inc = min(lrti_incidence, na.rm = TRUE),
            median_inc = median(lrti_incidence, na.rm = TRUE),
            max_inc = max(lrti_incidence, na.rm = TRUE))
incidence_summary

count_summary <- analysis_data %>% 
  group_by(treatment) %>% 
  filter(lrti_incidence < 0.7) %>% 
  summarize(mean_events = mean(lrti_events, na.rm = TRUE),
            sd_events = sd(lrti_events, na.rm = TRUE),
            n_events = n(),
            min_events = min(lrti_events, na.rm = TRUE),
            median_events = median(lrti_events, na.rm = TRUE),
            max_events = max(lrti_events, na.rm = TRUE))
count_summary

# Histogram for distribution of data
events_plot <- analysis_data %>% 
  ggplot() +
    geom_histogram(aes(lrti_events)) +
    theme_light() +
    facet_wrap(~treatment)
events_plot

events_di_plot <- analysis_data %>% 
  ggplot() +
    geom_boxplot(aes(lrti_events_di, person_time_at_risk)) +
    facet_wrap(~treatment)
events_di_plot

@@

nb_results <- glm.nb(lrti_events ~ treatment + offset(log_person_time_at_risk), data = analysis_data)
odTest(nb_results)

@@

# Link for analysis ideas and notes:
# https://stats.idre.ucla.edu/r/dae/negative-binomial-regression/ 
# https://stats.idre.ucla.edu/r/dae/poisson-regression/ 
# https://stats.idre.ucla.edu/r/faq/random-coefficient-poisson-models/

# Other diagnostic plots and tests
library(vcd)
fit <- goodfit(analysis_data$lrti_events)
summary(fit) 
rootogram(fit)
Ord_plot(analysis_data$lrti_events)

library(car)
influencePlot(nb_results)

library(pscl)

```

```{r}
# Mixed model variations

poisson_results <- glmer(lrti_events ~ treatment + child_age + (1 | home:cohort:area), 
                         data = analysis_data, family = poisson(link = "log"), offset = log_person_time_at_risk)
summary(poisson_results)

negbin_results <- glmer.nb(lrti_events ~ treatment + child_age + (1 | home:cohort:area) + offset(log_person_time_at_risk), 
                           data = analysis_data)
summary(negbin_results)

plot(negbin_results)
qqnorm(residuals(negbin_results))
qqline(residuals(negbin_results))

logreg_results <- glmer(lrti_events_di ~ treatment + child_age + (1 | home:cohort:area), 
                        data = analysis_data, family = binomial, offset = log_person_time_at_risk)
summary(logreg_results)
```
